Profit Optimisation in Heterogeneous 2-link Parallel Games

### 2025-02-13 thesis/article.tex
Thesis: Final fixes and checks, corrects Th. Lianea title to External Collaborator
Release v1.3

\renewcommand{\maketitle}{
	\begin{titlepage}
		\newgeometry{left=4cm, top=2.3cm, bottom=2.3cm}
		\hbox{\mbox{\hspace{-1.3cm}}
			\vrule depth 0.98\textheight
			\mbox{\hspace{1cm}}
			\vtop{
				\vspace{2cm}
				\begin{flushleft}
					\huge{\bf Pricing Games in Heterogeneous Parallel Networks\\}
					\vskip3cm
					\Large  Thomas Pappas\\
					AL1.18.0011\\
					\vskip3cm
					\begin{minipage}{7.5cm}
						\begin{flushleft}
							\normalsize {\it {\bf Examination committee:}\\
								Thanassis Lianeas,\\ School of Electrical and Computer Engineering, National Technical University of Athens.\\
								Dimitris Fotakis,\\ School of Electrical and Computer Engineering, National Technical University of Athens.\\
								Aris Pagourtzis,\\ School of Electrical and Computer Engineering, National Technical University of Athens.}
						\end{flushleft}
					\end{minipage}
					\hskip0.5cm
					\begin{minipage}{6cm}
						\begin{flushleft}
							\normalsize {\it {\bf Supervisor:}\\
								Thanassis Lianeas, Lecturer,\\School of Electrical and Computer Engineering,\\
								National Technical University of Athens.\\
							}
						\end{flushleft}
					\end{minipage}
				\end{flushleft}
				\vskip4cm
				\hskip4.5cm
				\begin{minipage}{5cm}
					\begin{center}
						\includegraphics[width=0.8\textwidth]{alma.png}
					\end{center}
		\end{minipage}}}
	\end{titlepage}
}

@@

\chapter*{Συνοψη}
Σε αυτή τη διπλωματική μεταπτυχιακού μελετάμε ένα παίγνιο $2$ επιπέδων όπου, σε ένα δίκτυο με $n$ ακμές από μια πηγή $s$ σε ένα στόχο $t$, μια μονάδα ροής θέλει να μετακινηθεί από το $s$ στο $t$ χρησιμοποιώντας την ακμή με το χαμηλότερο κόστος, ενώ οι διαχειριστές των ακμών ανταγωνίζονται για κέρδος βάζοντας δίοδια στις ακμές και δημιουργώντας έτσι ένα παίγνιο συμφόρησης με διόδια.
Εξετάζουμε μόνο αφινικές καθυστερήσεις για τις ακμές.
Επιπλέον, ο κάθε χρήστης δικτύου αντιδρά στα διόδια με ετερογενές τρόπο, δηλ. ο κάθε παίκτης $p$ στη ροή του δικτύου έχει διαφορετική τιμή ευαισθησίας χρόνου-χρήματος, η οποία περιγράφεται από μια συνάρτηση κατανομής $\alpha(p)$.
Πρώτα παρουσιάζουμε κάποιους βασικούς ορισμούς και ιδιότητες των ομοιογενών παιγνίων κέρδους, τα οποία είναι κυρίως βασισμένα σε προηγούμενες δουλειές, και με αυτά ως βάση, τα επεκτείνουμε για να περιγράψουμε παίγνια κέρδους με ετερογενείς χρήστες.
Εισαγάγουμε έναν νέο όρο, τη συνάρτηση διαχωρισμού ευαισθησίας $\as(t)$, ορισμένη για οποιεσδήποτε $2$ ακμές του δικτύου ως η τιμή ευαισθησίας χρόνου-χρήματος για την οποία, δωσμένων σετ διοδίων $t$, τα κόστη των $2$ ακμών είναι ίσα.
Με τη βοήθεια της $\as$ θα περιγράψουμε μερικές ιδιότητες γύρω από την εγωιστική συμπεριφορά των ετερογενών χρηστών για διαφορετικά διόδια, καθώς και θα αποδείξουμε την ύπαρξη άνω φράγματος για τις τιμές ευαισθησίας που αφορούν το παίγνιο.
Στη συνέχεια, φτάνοντας στο κύριο σώμα αυτής της έρευνας, εξετάζουμε την ύπαρξη ισορροπίας Nash στο παίγνιο κέρδους μεταξύ των διαχειριστών των ακμών για διαφορετικές περι-πτώσεις ετερογενών χρηστών.
Δείχνουμε ότι δεν υπάρχει ισορροπία Nash όταν υπάρ-χουν πολλοί χρήστες με $0$ ευαισθησία, δηλ. παίκτες που δεν επηρρεάζονται από τα διόδια, κάτι το οποίο μας αναγκάζει στο εξής να υποθέσουμε ότι $\alpha(p) > 0$.
Στη συνέχεια, εξετάζουμε περιπτώσεις όπου η συνάρτηση κατανομής είναι σταθερή και άρα η συμπεριφορά των χρηστών παραμένει ομογενής (κάνοντάς τα ψευδο-ετερογενή παίγνια), μέσω των οποίων μετά θα περιγράψουμε πώς συμπεριφέρονται η ροή, τα διόδια και τα σημεία ισορροπίας όταν αυτή η σταθερή τιμή μεταβάλλεται.
Χρησιμο-ποιώντας όλα τα παραπάνω, μελετάμε βηματικές συναρτήσεις κατανομής, όπου, περιο-ρίζοντας το μοντέλο μας σε στιγμιότυπα $2$ ακμών, αποδεικνύουμε μια ισχυρή απαίτηση για την τιμή ενός πιθανού σημείου ισορροπίας Nash στο παίγνιο κέρδους μεταξύ των διαχειριστών των ακμών, συνδέοντάς το με το σημείο ισορροπίας σε ένα ισοδύναμο ψευδο-ετερογενές παίγνιο.
Εντέλει, παρουσιάζουμε και αναλύουμε έναν αλγόριθμο για τον υπολογισμό αυτού του σημείου ισορροπίας, αν υπάρχει, σε χρόνο γραμμικό σε σχέση με τον πληθάριθμο της βηματικής συνάρτησης, και κλείνουμε με μια συζήτηση για πιθανές μελλοντικές εργασίες.

@@

Regarding the model properties, there is a multitude of variations in the classical selfish routing model.
Beginning with the network topology, one can distinguish between parallel networks (our case) or series-parallel ones.
In addition the flow traversing through the network can either be atomic (finite number of players, each contributing to congestion equally or weighted) or non-atomic (infinite number of players, each contributing to congestion by an non-influencing infinitesimal amount).
Also the model might admit elastic demand (e.g. \cite{10.1287/moor.1060.0231}, \cite{Hearn1998}), a property where users have a bound on the cost they're willing to pay, making a player potentially not participate in the flow if all link costs are above that bound.

Most work considers homogeneous network users.
Heterogeneity has been studied by Schmeilder \cite{1973JSP.....7..295S} while Milchtaich \cite[Prop $3.3$]{doi:10.1287/moor.25.3.349.12220} has contributed through with work on the more general crowding games, a class of games where the cost assigned to each player is affected only by the number of players selecting the same action (or strategy).
Large crowding games are basically less restrictive $n$-link parallel congestion games.
Also, Cole et al. \cite{10.1145/780542.780618} in their work showed that even general heterogeneous networks can be priced so that an optimal routing emerges while computing those efficient tolls can be done in polynomial time for convex latency functions and distribution functions with only finitely many values.
The latter paper has been one of the main pillars of this thesis.

@@

In this section we will describe the model utilised throughout this thesis.
If we attempted to call that model by including all of its properties, the result would be a Heterogeneous Non-atomic $n$-link Parallel Network Toll Congestion Pricing Game, so while we will describe each of those properties, we will eventually focus on heterogeneity and the pricing competition, thus calling it a Heterogeneous Pricing Game.

@@

We consider a directed graph $G = (\{s, t\}, N)$ where $N = \{1,\dots, n\}$ is a set of parallel links from a source node $s$ to a target node $t$.
For the non-atomic case, there is one unit of traffic that wishes to travel from $s$ to $t$, described as the unit interval $[0, 1]$ endowed with Lebesque measure $\lambda$.
Each point $p \in [0, 1]$ will be called a \textit{player} and will be considered to be non-cooperative and contribute to the traffic by an infinitesimal amount.
As such the decisions of individual players have no effect on the game and it becomes natural to only consider non-zero measure collection of players.

@@

In the homogeneous case all players have an equal reaction to tolls, and therefore the effective cost of a player using link $i$ becomes $\ell_i(x_i) + t_i$.
For a given set of tolls $t$, a flow $x$ is a \textit{Wardrop equilibrium for $t$} if $\forall i, j \in N$ with $x_i > 0$ it holds that
\begin{equation*}
	\ell_i(x_i) + t_i \leq \ell_j(x_j) + t_j
\end{equation*}
In that case, all links with $x_i > 0$ have equal effective costs, i.e. there exists some $K > 0$ such that for all those links $i$ it holds that $\ell_i(x_1) + t_i = K$.
From a well-known result from Beckman et al. \cite{beckmann1956studies} and Dafermos and Sparrow \cite{1363388843888284416}, such and an equilibrium for $t$ exists, is unique with regard to costs and can be described by the following inequality.
\begin{lemma}
	\label{lemma:wardrop_equilibrium}
	A flow $x$ is a Wardrop equilibrium for $t$ if and only if for all feasible flows $x^\prime$,
	\[\sum_{i \in N} (\ell_i(x_i) + t_i) \cdot (x_i - x_i^\prime) \leq 0\]
\end{lemma}

@@

We can therefore denote by $x(t)$ the flow occurring on that unique up to costs equilibrium for a given set of toll $t$.
If $t = 0^N$ then the equilibrium is called the \textit{Wardrop equilibrium}.
It's also worth noting that looking at Lemma \ref{lemma:wardrop_equilibrium}, we can see that what actually matters, with regard to the flow, is the relative difference among the tolls, and not their actual values.
To demonstrate this, if we consider an arbitrary $c \in \R_+$ and tolls $t^\prime = t + c$ then for the same flow $x$ and any other flow $x^\prime$ the sum becomes
\begin{flalign*}
	\sum_{i \in N} (\ell_i(x_i) + t_i + c) (x_i - x_i^\prime) &= \sum_{i \in N} (\ell_i(x_i) + t_i) (x_i - x_i^\prime) + c \left(\sum_{i \in N} x_i - \sum_{i \in N} x_i^\prime\right) \\
	&= \sum_{i \in N} (\ell_i(x_i) + t_i) (x_i - x_i^\prime) \leq 0
\end{flalign*}
which makes $x$ the Wardrop equilibrium for $t^\prime$ as well.
The last equality holds because $x_i, x_i^\prime$ are flows of the network and therefore $\sum_{i \in N} x_i = \sum_{i \in N} x_i^\prime = 1$.
Also, this reasoning is still valid for $c < 0$ as long as $t_i + c \ge 0$ for all $i \in N$, or in other words, as long as $c \ge -\min_{i \in N}{t_i}$.

With this observation we can see that the case of $t = 0^N$ is equivalent to any toll instance such that $t_1 = t_2 = \dots = t_n$.
We will therefore simplify our notation and write $t = 0$ when talking about the game without tolls, though we will in fact mean the entire class $\{t = (c)_{i \in N}, c \in \R_+\}$.

Finally for affine latencies we can calculate $x(t)$ by following the same reasoning as \cite[Prop $3.1$]{Harks_2019}.
For $\ell_i(x_i) = a_i x_i + b_i$ with $a_i > 0, b_i \geq 0$ and $t \in \R_+^{|N|}$, define $N(t) = \{i \in N | x_i(t) > 0\}$.
Since $x(t)$ is an equilibrium for $t$ then for all $i \in N(t)$ we get $\ell_i(x_i) + t_i = K$ for some common effective cost $K$.
By also using the fact that $\sum_{j \in N(t)} x_i = 1$ we solve the equations with regard to $K$ and eventually get for all $i \in N(t)$
\begin{equation}
	\label{eq:homogeneous_x_i}
	x_i(t) = \frac{1 + \sum_{j \in N(t)}\frac{b_j + t_j - b_i - t_i}{a_j}}{\sum_{j \in N(t)}\frac{a_i}{a_j}}
\end{equation}

@@

This property shows us that for a given $t \in \R_+^{|N|}$, if we order the links according to decreasing toll values, then the same link order will have latencies ordered in increasing values (and vice-versa).
It is then easy to show that the players using each such link are also ordered with regard to their $\alpha$ values in increasing order.
Therefore a given set of tolls can define an ordering of links where $t_1 \ge t_2 \ge \dots \ge t_n$ and $\ell_1(x_1(t)) \le \ell_2(x_2(t)) \le \dots \le \ell_n(x_n(t))$, while if we were to pick player representatives from each link with positive flow $p_1, p_2, \dots, p_k$, we also get $\alpha(p_1) \le \alpha(p_2) \le \dots \le \alpha(p_k)$.
Note that the ordering is unique only when the inequalities are strict, as had we had $t_i = t_j$ for any $i, j \in N$ then $i$ and $j$ can be ordered arbitrarily.
If $t = 0$ then all links can be ordered arbitrarily, which is why when comparing $t$ with $0$ we will assume the ordering is the same for simplicity.
We continue with the following lemma which describes a property for the last link in such an ordering.

\begin{lemma}
	\label{lemma:xn_xn0_lower_bound}
	Let $(N, \ell, \alpha)$ a heterogeneous parallel game.
	For tolls $t \in \R_+^{|N|}$ consider the links in decreasing toll ordering.
	It holds that $x_n(t) \ge x_n(0)$.
\end{lemma}

\begin{proof}
	Consider $t^\prime = t - t_n$.
	From assumption we have $t_n = \min_{i \in N} t_i$, which makes $t^\prime \in \R_+^{|N|}$ with $t_n^\prime = 0$.
	Since the flow equilibrium $x(t)$ is dependent only on the toll differences, we get $x(t) = x(t^\prime)$.
	Also, by now comparing $t^\prime$ with $t = 0$, no additional tolls have been added to link $n$, making $x_n(t^\prime) \ge x_n(0)$.
	The proposition follows naturally from the previous two statements.
\end{proof}

Finally we need to discuss about toll differences again.
While the definition for $t \in \R_+^{|N|}$ allows any arbitrary value, the meaningful ones actually reside in a more limited space.
Assume for example that a toll $t$ is such that $x_i(t) = 1$ for some link $i \in N$ with $t_i > 0$ and $x_j(t) = 0$ for all other links $j \ne i$.
It's obvious that further decreasing $t_i$ changes nothing for the flow, which is why in any optimisation setup (minimum total cost, maximal profit, etc.) those cases never come into play. Same reasoning can be done for increasing $t_i$ in tolls $t$ where $x_i(t) = 0$.
Therefore there exists an upper bound for the valid toll difference between any two links $i, j \in N$, which can be calculated as the toll difference which even the least sensitive player $0$ (to account for heterogeneity) will view the link with the lowest toll (and all flow) as the one with the lowest cost.
In other words, assuming $t_i > t_j$ with all flow in $j$ and $\alpha(0) > 0$ we get
\[\ell_j(1) + \alpha(0) \cdot t_i \le \ell_i(0) + \alpha(0) \cdot t_j \Rightarrow t_i - t_j \le \frac{\ell_j(1) - \ell_i(0)}{\alpha(0)}\]
In the following chapters we will generalise and investigate more formally this idea, along with the overall relation between the toll and latency differences with regard to sensitivity values.

@@

We can easily calculate that $x(0) = (2/3, 1/3)$ with equal cost $4/3$ in both links.
If tolls $t = (3, 2)$ are assigned to the links, then $x(3, 2) = (1/4, 3/4)$.
Indeed, since $t_1 > t_2$ then the flow users with the lowest sensitivity will use link $1$, while the more sensitive ones will use link $2$.
With players ordered in increasing sensitivity order, the less sensitive link $1$ users will have players $p_1 \in [0, 1/4]$ with $\alpha(p_1) \le \alpha(1/4) = 5/4$ and respectively link $2$ users $p_2 \in (1/4, 1]$ (player $x_i$ can be put arbitrarily in link $1$ or $2$) we have $\alpha(p_2) \ge \alpha(1/4) = 5/4$.
Finally, all flow users will view the link latencies similarly as
\[
\ell\left(\tfrac14, \tfrac34\right) = \left(\ell_1\left(\tfrac14\right), \ell_2\left(\tfrac34\right)\right) = \left(\tfrac24, \tfrac74\right)
\]
With all the above we can verify now that for all players in link $1$ it holds that
\[
c_1(p) \le c_2(p) \Leftrightarrow \tfrac24 + 3\alpha(p_1) \le \tfrac74 + 2\alpha(p) \Leftrightarrow \alpha(p) \le \tfrac54
\]
while the opposite direction also holds similarly for players in link $2$.
We postpone the discussion of how that flow can be calculated until Chapter \ref{chapter:split}, as we'll require a better notion of the player who views the two links as equal (in this case player $1/4$).

@@

From the work of Harkes et al. \cite[Example 4.2]{Harks_2019}, we know that, for homogeneous players, uniqueness of $t^*$ is guaranteed only when $x_i(0) > 0$ for all $i \in N$, a property called the \textit{full Wardrop support assumption}.
This is seemingly not as much restrictive in practice, as there would be little motivation to add tolls to a route with no traffic.
From the same work we also get that if the full Wardrop support assumption holds, then $t^*$ is unique \cite[Lemma 3.3]{Harks_2019} while also $x_i(t^*) > 0$ and $t_i^* > 0$ for all $i \in N$ \cite[Lemma 3.2]{Harks_2019}.
%We will thus assume for all further instances that the full Wardrop support assumption is true.
Finally, again from Harkes et al. \cite[Lemma 3.3]{Harks_2019} we can also get a formula for the Nash Equilibrium of the pricing game for affine latencies $\ell_i(x) = a_i x + b_i$.
\begin{equation}
	\label{eq:homogeneous_pricing_ne_i}
	t_i^* = \left(a_i + \frac{1}{\sum_{j \ne i} \frac{1}{a_j}}\right) \cdot x_i(t^*)
\end{equation}

@@

One of the main difficulties with heterogeneity is that, for a given set of tolls, flow users with different money sensitivity values, view toll values also differently, and therefore, in practice, play different games.
In the analysis of Example \ref{example:simple_alpha} we can find a first glimpse on how to tackle this issue, as we can see that the flow users on each link are organised around the player that views the two link costs as equal.
This notion will be the center of this chapter, and using it we will try and describe some properties about all users.

>>>>

\renewcommand{\maketitle}{
	\begin{titlepage}
		\newgeometry{left=4cm, top=2.3cm, bottom=2.3cm}
		\hbox{\mbox{\hspace{-1.3cm}}
			\vrule depth 0.98\textheight
			\mbox{\hspace{1cm}}
			\vtop{
				\vspace{2cm}
				\begin{flushleft}
					\huge{\bf Pricing Games in Heterogeneous Parallel Networks\\}
					\vskip3cm
					\Large  Thomas Pappas\\
					AL1.18.0011\\
					\vskip3cm
					\begin{minipage}{7.5cm}
						\begin{flushleft}
							\normalsize {\it {\bf Examination committee:}\\
								Thanasis Lianeas,\\ External Collaborator, ALMA.\\
								Dimitris Fotakis,\\ School of Electrical and Computer Engineering, National Technical University of Athens.\\
								Aris Pagourtzis,\\ School of Electrical and Computer Engineering, National Technical University of Athens.}
						\end{flushleft}
					\end{minipage}
					\hskip0.5cm
					\begin{minipage}{6cm}
						\begin{flushleft}
							\normalsize {\it {\bf Supervisor:}\\
								Thanasis Lianeas,\\ External Collaborator, ALMA.\\
							}
						\end{flushleft}
					\end{minipage}
				\end{flushleft}
				\vskip4cm
				\hskip4.5cm
				\begin{minipage}{5cm}
					\begin{center}
						\includegraphics[width=0.8\textwidth]{alma.png}
					\end{center}
		\end{minipage}}}
	\end{titlepage}
}

@@

\chapter*{Συνοψη}
Σε αυτή τη διπλωματική εργασία μεταπτυχιακού μελετάμε ένα παίγνιο $2$ επιπέδων όπου, σε ένα δίκτυο με $n$ ακμές από μια πηγή $s$ σε ένα στόχο $t$, μια μονάδα ροής θέλει να μετακινηθεί από το $s$ στο $t$ χρησιμοποιώντας την ακμή με το χαμηλότερο κόστος, ενώ οι διαχειριστές των ακμών ανταγωνίζονται για κέρδος βάζοντας δίοδια στις ακμές και δημιουργώντας έτσι ένα παίγνιο συμφόρησης με διόδια.
Εξετάζουμε μόνο αφινικές καθυστερήσεις για τις ακμές.
Επιπλέον, ο κάθε χρήστης δικτύου αντιδρά στα διόδια με ετερογενές τρόπο, δηλ. ο κάθε παίκτης $p$ στη ροή του δικτύου έχει διαφορετική τιμή ευαισθησίας χρόνου-χρήματος, η οποία περιγράφεται από μια συνάρτηση κατανομής $\alpha(p)$.
Πρώτα παρουσιάζουμε κάποιους βασικούς ορισμούς και ιδιότητες των ομοιογενών παιγνίων κέρδους, τα οποία είναι κυρίως βασισμένα σε προηγούμενες δουλειές, και με αυτά ως βάση, τα επεκτείνουμε για να περιγράψουμε παίγνια κέρδους με ετερογενείς χρήστες.
Εισαγάγουμε έναν νέο όρο, τη συνάρτηση διαχωρισμού ευαισθησίας $\as(t)$, ορισμένη για οποιεσδήποτε $2$ ακμές του δικτύου ως η τιμή ευαισθησίας χρόνου-χρήματος για την οποία, δωσμένων σετ διοδίων $t$, τα κόστη των $2$ ακμών είναι ίσα.
Με τη βοήθεια της $\as$ θα περιγράψουμε μερικές ιδιότητες γύρω από την εγωιστική συμπεριφορά των ετερογενών χρηστών για διαφορετικά διόδια, καθώς και θα αποδείξουμε την ύπαρξη άνω φράγματος για τις τιμές ευαισθησίας που αφορούν το παίγνιο.
Στη συνέχεια, φτάνοντας στο κύριο σώμα αυτής της έρευνας, εξετάζουμε την ύπαρξη ισορροπίας Nash στο παίγνιο κέρδους μεταξύ των διαχειριστών των ακμών για διαφορετικές περι-πτώσεις ετερογενών χρηστών.
Δείχνουμε ότι δεν υπάρχει ισορροπία Nash όταν υπάρ-χουν πολλοί χρήστες με $0$ ευαισθησία, δηλ. παίκτες που δεν επηρρεάζονται από τα διόδια, κάτι το οποίο μας αναγκάζει στο εξής να υποθέσουμε ότι $\alpha(p) > 0$.
Στη συνέχεια, εξετάζουμε περιπτώσεις όπου η συνάρτηση κατανομής είναι σταθερή και άρα η συμπεριφορά των χρηστών παραμένει ομογενής (κάνοντάς τα ψευδο-ετερογενή παίγνια), μέσω των οποίων μετά θα περιγράψουμε πώς συμπεριφέρονται η ροή, τα διόδια και τα σημεία ισορροπίας όταν αυτή η σταθερή τιμή μεταβάλλεται.
Χρησιμο-ποιώντας όλα τα παραπάνω, μελετάμε βηματικές συναρτήσεις κατανομής, όπου, περιο-ρίζοντας το μοντέλο μας σε στιγμιότυπα $2$ ακμών, αποδεικνύουμε μια ισχυρή απαίτηση για την τιμή ενός πιθανού σημείου ισορροπίας Nash στο παίγνιο κέρδους μεταξύ των διαχειριστών των ακμών, συνδέοντάς το με το σημείο ισορροπίας σε ένα ισοδύναμο ψευδο-ετερογενές παίγνιο.
Εντέλει, παρουσιάζουμε και αναλύουμε έναν αλγόριθμο για τον υπολογισμό αυτού του σημείου ισορροπίας, αν υπάρχει, σε χρόνο γραμμικό σε σχέση με τον πληθάριθμο της βηματικής συνάρτησης, και κλείνουμε με μια συζήτηση για πιθανές μελλοντικές εργασίες.

@@

Regarding the model properties, there is a multitude of variations in the classical selfish routing model.
Beginning with the network topology, one can distinguish between parallel networks (our case) or series-parallel ones.
In addition, the flow traversing through the network can either be atomic (finite number of players, each contributing to congestion equally or weighted) or non-atomic (infinite number of players, each contributing to congestion by an non-influencing infinitesimal amount).
Also the model might admit elastic demand (e.g. \cite{10.1287/moor.1060.0231}, \cite{Hearn1998}), a property where users have a bound on the cost they're willing to pay, making a player potentially not participate in the flow if all link costs are above that bound.

Most work considers homogeneous network users.
Heterogeneity has been studied by Schmeilder \cite{1973JSP.....7..295S} while Milchtaich \cite[Prop $3.3$]{doi:10.1287/moor.25.3.349.12220} has contributed through his work on the more general crowding games, a class of games where the cost assigned to each player is affected only by the number of players selecting the same action (or strategy).
Large crowding games are basically less restrictive $n$-link parallel congestion games.
Also, Cole et al. \cite{10.1145/780542.780618}, in their work, showed that even general heterogeneous networks can be priced so that an optimal routing emerges and computing those efficient tolls can be done in polynomial time for convex latency functions and distribution functions with only finitely many values.
The latter paper has been one of the main pillars of this thesis.

@@

In this chapter we will describe the model utilised throughout this thesis.
If we attempted to call that model by including all of its properties, the result would be a Heterogeneous Non-atomic $n$-link Parallel Network Toll Congestion Pricing Game, so while we will describe each of those properties, we will eventually focus on heterogeneity and the pricing competition, thus calling it a Heterogeneous Pricing Game.

@@

We consider a directed graph $G = (\{s, t\}, N)$ where $N = \{1,\dots, n\}$ is a set of parallel links from a source node $s$ to a target node $t$.
For the non-atomic case, there is one unit of traffic that wishes to travel from $s$ to $t$, described as the unit interval $[0, 1]$, endowed with Lebesque measure $\lambda$.
Each point $p \in [0, 1]$ will be called a \textit{player} and will be considered to be non-cooperative and contribute to the traffic by an infinitesimal amount.
As such the decisions of individual players have no effect on the game and it becomes natural to only consider non-zero measure collection of players.

@@

In the homogeneous case all players have an equal reaction to tolls, and therefore the effective cost of a player using link $i$ becomes $\ell_i(x_i) + t_i$.
For a given set of tolls $t$, a flow $x$ is a \textit{Wardrop equilibrium for $t$} if $\forall i, j \in N$ with $x_i > 0$ it holds that
\begin{equation*}
	\ell_i(x_i) + t_i \leq \ell_j(x_j) + t_j
\end{equation*}
In that case, all links with $x_i > 0$ have equal effective costs, i.e. there exists some $K > 0$ such that for all those links $i$ it holds that $\ell_i(x_i) + t_i = K$.
From a well-known result from Beckman et al. \cite{beckmann1956studies} and Dafermos and Sparrow \cite{1363388843888284416}, such and an equilibrium for $t$ exists, is unique with regard to costs and can be described by the following inequality.
\begin{lemma}
	\label{lemma:wardrop_equilibrium}
	A flow $x$ is a Wardrop equilibrium for $t$ if and only if for all feasible flows $x^\prime$,
	\[\sum_{i \in N} (\ell_i(x_i) + t_i) \cdot (x_i - x_i^\prime) \leq 0\]
\end{lemma}

@@

We can therefore denote by $x(t)$ the flow occurring on that unique -up to costs- equilibrium for a given set of toll $t$.
If $t = 0^n$ then the equilibrium is called the \textit{Wardrop equilibrium}.
It's also worth noting that, looking at Lemma \ref{lemma:wardrop_equilibrium}, we can see that what actually matters, with regard to the flow, is the relative difference among the tolls and not their actual values.
To demonstrate this, if we consider an arbitrary $c \in \R_+$ and tolls $t^\prime = t + c$ then for the same flow $x$ and any other flow $x^\prime$ the sum becomes
\begin{flalign*}
	\sum_{i \in N} (\ell_i(x_i) + t_i + c) (x_i - x_i^\prime) &= \sum_{i \in N} (\ell_i(x_i) + t_i) (x_i - x_i^\prime) + c \left(\sum_{i \in N} x_i - \sum_{i \in N} x_i^\prime\right) \\
	&= \sum_{i \in N} (\ell_i(x_i) + t_i) (x_i - x_i^\prime) \leq 0
\end{flalign*}
which makes $x$ the Wardrop equilibrium for $t^\prime$ as well.
The last equality holds because $x_i, x_i^\prime$ are flows of the network and therefore $\sum_{i \in N} x_i = \sum_{i \in N} x_i^\prime = 1$.
Also, this reasoning is still valid for $c < 0$ as long as $t_i + c \ge 0$ for all $i \in N$, or in other words, as long as $c \ge -\min_{i \in N}{t_i}$.

With this observation we can see that the case of $t = 0^n$ is equivalent to any toll instance such that $t_1 = t_2 = \dots = t_n$.
We will therefore simplify our notation and write $t = 0$ when talking about the game without tolls, though we will in fact mean the entire class $\{t = (c)_{i \in N}, c \in \R_+\}$.

Finally, for affine latencies we can calculate $x(t)$ by following the same reasoning as \cite[Prop $3.1$]{Harks_2019}.
For $\ell_i(x_i) = a_i x_i + b_i$ with $a_i > 0, b_i \geq 0$ and $t \in \R_+^n$, define $N(t) = \{i \in N | x_i(t) > 0\}$.
Since $x(t)$ is an equilibrium for $t$ then for all $i \in N(t)$ we get $\ell_i(x_i) + t_i = K$ for some common effective cost $K$.
By also using the fact that $\sum_{j \in N(t)} x_i = 1$ we solve the equations with regard to $K$ and eventually get for all $i \in N(t)$
\begin{equation}
	\label{eq:homogeneous_x_i}
	x_i(t) = \frac{1 + \sum_{j \in N(t)}\frac{b_j + t_j - b_i - t_i}{a_j}}{\sum_{j \in N(t)}\frac{a_i}{a_j}}
\end{equation}

@@

This property shows us that for a given $t \in \R_+^n$, if we order the links according to decreasing toll values, then the same link order will have latencies ordered in increasing values (and vice-versa).
It is then easy to show that the players using each such link are also ordered with regard to their $\alpha$ values in increasing order.
Therefore a given set of tolls can define an ordering of links where $t_1 \ge t_2 \ge \dots \ge t_n$ and $\ell_1(x_1(t)) \le \ell_2(x_2(t)) \le \dots \le \ell_n(x_n(t))$, while if we were to pick player representatives from each link with positive flow $p_1, p_2, \dots, p_k$, we also get $\alpha(p_1) \le \alpha(p_2) \le \dots \le \alpha(p_k)$.
Note that the ordering is unique only when the inequalities are strict, as had we had $t_i = t_j$ for any $i, j \in N$ then $i$ and $j$ can be ordered arbitrarily.
If $t = 0$ then all links can be ordered arbitrarily, which is why when comparing $t$ with $0$ we will assume the ordering is the same for simplicity.
We continue with the following lemma which describes a property for the last link in such an ordering.

\begin{lemma}
	\label{lemma:xn_xn0_lower_bound}
	Let $(N, \ell, \alpha)$ a heterogeneous parallel game.
	For tolls $t \in \R_+^n$ consider the links in decreasing toll ordering.
	It holds that $x_n(t) \ge x_n(0)$.
\end{lemma}

\begin{proof}
	Consider $t^\prime = t - t_n$.
	From assumption we have $t_n = \min_{i \in N} t_i$, which makes $t^\prime \in \R_+^n$ with $t_n^\prime = 0$.
	Since the flow equilibrium $x(t)$ is dependent only on the toll differences, we get $x(t) = x(t^\prime)$.
	Also, by now comparing $t^\prime$ with $t = 0$, no additional tolls have been added to link $n$, making $x_n(t^\prime) \ge x_n(0)$.
	The proposition follows naturally from the previous two statements.
\end{proof}

Finally we need to discuss about toll differences again.
While the definition for $t \in \R_+^n$ allows any arbitrary value, the meaningful ones actually reside in a more limited space.
Assume, for example, that a toll $t$ is such that $x_i(t) = 1$ for some link $i \in N$ with $t_i > 0$ and $x_j(t) = 0$ for all other links $j \ne i$.
It's obvious that further decreasing $t_i$ changes nothing for the flow, which is why in any optimisation setup (minimum total cost, maximal profit, etc.) those cases never come into play. Same reasoning can be done for increasing $t_i$ in tolls $t$ where $x_i(t) = 0$.
Therefore, there exists an upper bound for the valid toll difference between any two links $i, j \in N$, which can be calculated as the toll difference which even the least sensitive player $0$ (to account for heterogeneity) will view the link with the lowest toll (and all flow) as the one with the lowest cost.
In other words, assuming $t_i > t_j$ with all flow in $j$ and $\alpha(0) > 0$ we get
\[\ell_j(1) + \alpha(0) \cdot t_i \le \ell_i(0) + \alpha(0) \cdot t_j \Rightarrow t_i - t_j \le \frac{\ell_j(1) - \ell_i(0)}{\alpha(0)}\]
In the following chapters we will generalise and investigate more formally this idea, along with the overall relation between the toll and latency differences with regard to sensitivity values.

@@

We can easily calculate $x(0) = (2/3, 1/3)$ with equal cost $4/3$ in both links.
If tolls $t = (3, 2)$ are assigned to the links, then $x(3, 2) = (1/4, 3/4)$.
Indeed, since $t_1 > t_2$ then the flow users with the lowest sensitivity will use link $1$, while the more sensitive ones will use link $2$.
With players ordered in increasing sensitivity order, the less sensitive link $1$ users will have players $p_1 \in [0, 1/4]$ with $\alpha(p_1) \le \alpha(1/4) = 5/4$ and respectively link $2$ users $p_2 \in (1/4, 1]$ (player $x_i$ can be put arbitrarily in link $1$ or $2$) have $\alpha(p_2) \ge \alpha(1/4) = 5/4$.
Finally, all flow users will view the link latencies similarly as
\[
\ell\left(\tfrac14, \tfrac34\right) = \left(\ell_1\left(\tfrac14\right), \ell_2\left(\tfrac34\right)\right) = \left(\tfrac24, \tfrac74\right)
\]
With all the above we can verify now that for all players in link $1$ it holds that
\[
c_1(p) \le c_2(p) \Leftrightarrow \tfrac24 + 3\alpha(p_1) \le \tfrac74 + 2\alpha(p) \Leftrightarrow \alpha(p) \le \tfrac54
\]
while the opposite direction also holds similarly for players in link $2$.
We postpone the discussion of how that flow can be calculated until Chapter \ref{chapter:split}, as we'll require a better notion of the player who views the two links as equal (in this case, player $1/4$).

@@

From the work of Harkes et al. \cite[Example 4.2]{Harks_2019}, we know that, for homogeneous players, uniqueness of $t^*$ is guaranteed only when $x_i(0) > 0$ for all $i \in N$, a property called the \textit{full Wardrop support assumption}.
This is seemingly not as much restrictive in practice, as there would be little motivation to add tolls to a route with no traffic.
From the same work we also get that if the full Wardrop support assumption holds, then $t^*$ is unique \cite[Lemma 3.3]{Harks_2019}, while also $x_i(t^*) > 0$ and $t_i^* > 0$ for all $i \in N$ \cite[Lemma 3.2]{Harks_2019}.
%We will thus assume for all further instances that the full Wardrop support assumption is true.
Finally, again from Harkes et al. \cite[Lemma 3.3]{Harks_2019} we can also get a formula for the Nash Equilibrium of the pricing game for affine latencies $\ell_i(x) = a_i x + b_i$.
\begin{equation}
	\label{eq:homogeneous_pricing_ne_i}
	t_i^* = \left(a_i + \frac{1}{\sum_{j \ne i} \frac{1}{a_j}}\right) \cdot x_i(t^*)
\end{equation}

@@

One of the main difficulties with heterogeneity is that, for a given set of tolls, flow users with different money sensitivity values, view toll values also differently, and therefore, in practice, play different games.
In the analysis of Example \ref{example:simple_alpha}, we can see a first glimpse on how to tackle this issue, as we found out that the flow users on each link are organised around the player that views the two link costs as equal.
This notion will be the center of this chapter, and using it we will try and describe some properties about all users.
